# axis

## axis.title

:::::: {.columns style="display: flex;"}
::: {.column width="49%"}
``` {r, axis.title.1, eval = FALSE}
p + theme(
    axis.title          = element_text(size = 20, margin = margin(1,1,1,1,"line"))
#   axis.title.x        = element_text(),
#   axis.title.x.top    = element_text(),
#   axis.title.x.bottom = element_text(),
#   axis.title.y        = element_text(),
#   axis.title.y.left   = element_text(),
#   axis.title.y.right  = element_text(),
  )
```
* you may find `axis.title.*`, `axis.title.*.*` do not inherit `element_text(margin)` from `axis.title`. 
* it's because that in [default theme](#ggplot2-default-themes), `axis.title.x` and `axis.title.y` have called `element_text(margin)` already
* and which result in a gap between `axis.title` and `axis.text` in size `unit(11 / 4, "pt")`.
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="49%"}
``` {r, echo = FALSE}
pl <- {
  <<axis.title.1>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::::::


:::::: {.columns style="display: flex;"}
::: {.column width="49%"}
``` {r, axis.title.2, eval = FALSE}
p + theme(
#   axis.title          = element_text(),
    axis.title.x        = element_text(size = 20, margin = margin(1,1,1,1,"line")),
#   axis.title.x.top    = element_text(),
#   axis.title.x.bottom = element_text(),
    axis.title.y        = element_text(size = 20, margin = margin(1,1,1,1,"line"))
#   axis.title.y.left   = element_text(),
#   axis.title.y.right  = element_text(),
  )
```
* also, you may find `axis.title.x.top` dose not inherit `element_text(margin)` from `axis.title.x` 
* and `axis.title.y.right` dose not inherit that from `axis.title.y`.
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="49%"}
``` {r, echo = FALSE}
pl <- {
  <<axis.title.2>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::::::


:::::: {.columns style="display: flex;"}
::: {.column width="49%"}
``` {r, axis.title.3, eval = FALSE}
p + theme(
    axis.title          = element_text(size = 20),
    axis.title.x        = element_text(margin = margin(1,1,1,1,"line")),
    axis.title.x.top    = element_text(margin = margin(1,1,1,1,"line")),
#   axis.title.x.bottom = element_text(),
    axis.title.y        = element_text(margin = margin(1,1,1,1,"line")),
#   axis.title.y.left   = element_text(),
    axis.title.y.right  = element_text(margin = margin(1,1,1,1,"line"))
  )
```
* so to change `element_text(margin)`, these `axis.title.**` have to be set explicitly 
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="49%"}
``` {r, echo = FALSE}
pl <- {
  <<axis.title.3>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::::::


## axis.text

:::::: {.columns style="display: flex;"}
::: {.column width="49%"}
```{r axis.text.1, eval = FALSE}
p + theme(
    axis.text = element_text(size = 20, margin = margin(1,1,1,1,"line"))
#   axis.text.x        = element_text(),
#   axis.text.x.top    = element_text(),
#   axis.text.x.bottom = element_text(),
#   axis.text.y        = element_text(),
#   axis.text.y.left   = element_text(),
#   axis.text.y.right  = element_text()
  )
```
* same as `axis.title`
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="49%"}
``` {r, echo = FALSE}
pl <- {
  <<axis.text.1>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::::::

:::::: {.columns style="display: flex;"}
::: {.column width="49%"}
```{r axis.text.2, eval = FALSE}
p + theme(
    axis.text = element_text(size = 20),
#   axis.text.x        = element_text(),
    axis.text.x.top    = element_text(vjust = 1),
    axis.text.x.bottom = element_text(vjust = 0),
#   axis.text.y        = element_text(),
    axis.text.y.left   = element_text(hjust = 0),
    axis.text.y.right  = element_text(hjust = 1)
  )
```
* and there are also gaps between `axis.text` and `axit.ticks`, which due to [default theme](#ggplot2-default-themes) with default size `unit(0.8 * 11 / 4, "pt")`.
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="49%"}
```{r, echo = FALSE}
pl <- {
  <<axis.text.2>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE}
pl
```
```{r, echo = FALSE, include = FALSE, eval=F}
w.pt <- grid::convertUnit(unit(wh[1], units = "in"), "bigpts", valueOnly = T)
h.pt <- grid::convertUnit(unit(wh[2], units = "in"), "bigpts", valueOnly = T)
dpi <- 300
w.px <- w.pt * dpi / 72
h.px <- h.pt * dpi / 72
geometry <- paste(0.3 * w.px, "x", 0.25 * h.px, "+", 0.03 * w.px, "+", 0.05 * h.px)

fig <- magick::image_graph(width = w.px, height = h.px, res = dpi)
pl
dev.off()
```
```{r, echo = F}
#fig <- magick::image_crop(fig, geometry)
#cowplot::ggdraw() + cowplot::draw_image(fig)
cut_a_plot(pl, wh, width = c(0.03, 0.3), height = c(0.05, 0.25))
#rm(w.pt, h.pt, dpi, w.px, h.px, fig, geometry)
```
:::
::::::

## axis.ticks
`axis.ticks = element_line(size)` only controls the length along axes, and does not effect the size of plot.

:::: {.columns style="display: flex;"}
::: {.column width="49%"}
```{r, axis.ticks.1.1, eval = FALSE}
p + theme(
    axis.ticks = element_line(size = 10)
#   axis.ticks.x        = element_line(),
#   axis.ticks.x.top    = element_line(),
#   axis.ticks.x.bottom = element_line(),
#   axis.ticks.y        = element_line(),
#   axis.ticks.y.left   = element_line(),
#   axis.ticks.y.right  = element_line()
  )
```
``` {r, echo = FALSE}
pl <- {
  <<axis.ticks.1.1>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="49%"}
```{r, axis.ticks.1.2, eval = FALSE}
p + theme(
    axis.ticks = element_line(size = 100)
#   axis.ticks.x        = element_line(),
#   axis.ticks.x.top    = element_line(),
#   axis.ticks.x.bottom = element_line(),
#   axis.ticks.y        = element_line(),
#   axis.ticks.y.left   = element_line(),
#   axis.ticks.y.right  = element_line()
  )
```
``` {r, echo = FALSE}
pl <- {
  <<axis.ticks.1.2>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::::

To control the length vertically, you need to change `axis.ticks.length`.

:::: {.columns style="display: flex;"}
::: {.column width="49%"}
```{r, axis.ticks.2, eval = FALSE}
p + theme(
    axis.ticks.length = unit(0.5,"cm") # default unit(11 / 4, "pt")
#   axis.ticks.length.x        = NULL, 
#   axis.ticks.length.x.top    = NULL,
#   axis.ticks.length.x.bottom = NULL,
#   axis.ticks.length.y        = NULL,
#   axis.ticks.length.y.left   = NULL,
#   axis.ticks.length.y.right  = NULL
  )
```
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="49%"}
``` {r, echo = FALSE}
pl <- {
  <<axis.ticks.2>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::::

## axis.line
`axis.line` has no effect on plot size as well.

```{r}
gt1 <- ggplotGrob(p)
gt1$heights

p2 <- p + theme(
    axis.line = element_line(size = 10, linetype = 'dashed')
#   axis.line.x        = element_line(),
#   axis.line.x.top    = element_line(),
#   axis.line.x.bottom = element_line(),
#   axis.line.y        = element_line(),
#   axis.line.y.left   = element_line(),
#   axis.line.y.right  = element_line()
  )
gt2 <- ggplotGrob(p2)
gt2$heights
```
``` {r, echo = FALSE}
wh <- findWH(p2, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, fig.align = 'center', out.width = "60%"}
p2
```


## summary
```{r, echo = FALSE, fig.align='center'}
knitr::include_graphics(file.path(getwd(), "01.axis.summary.png"))
```
