# legend

```{r show_legend, echo = FALSE}
show_legend <- function(p, wh){
  w.pt <- grid::convertUnit(unit(wh[1], units = "in"), "bigpts", valueOnly = T)
  h.pt <- grid::convertUnit(unit(wh[2], units = "in"), "bigpts", valueOnly = T)
  dpi <- 300
  w.px <- w.pt * dpi / 72
  h.px <- h.pt * dpi / 72
  
  fig <- magick::image_graph(width = w.px, height = h.px, res = dpi)
  print(p)
  dev.off()
  
  geometry <- paste(1 * w.px, "x", 1 * h.px, "+", 0.66 * w.px, "+", 0 * h.px)
  
  fig <- magick::image_crop(fig, geometry)
  return(fig)
}
show_legend <- function(p, wh){
  cut_a_plot(p, wh, width = c(0.66, 1))
}
```

## legend.position
### in character {-}
Default `legend.position` is `right`, and also support `left`, `top`, `bottom` and `none`.

:::: {.columns style="display: flex;"}
::: {.column width="49%"}
```{r, legend.position.1, eval = FALSE}
p + theme(legend.position = "left")
```
``` {r, echo = FALSE}
pl <- {
  <<legend.position.1>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="49%"}
```{r, legend.position.3, eval = FALSE}
p + theme(legend.position = "top")
```
``` {r, echo = FALSE}
pl <- {
  <<legend.position.3>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::::

:::: {.columns style="display: flex;"}
::: {.column width="49%"}
```{r, legend.position.4, eval = FALSE}
p + theme(legend.position = "bottom")
```
``` {r, echo = FALSE}
pl <- {
  <<legend.position.4>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="49%"}
```{r, legend.position.2, eval = FALSE}
p + theme(legend.position = "none")
```
``` {r, echo = FALSE}
pl <- {
  <<legend.position.2>>
}
#wh <- findWH(pl, unit = "in", plot.size = unit(250, "bigpts"), is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::::

### in numeric vector {-}
`legend.position` also support a coordinate (two-element numeric vector).

A better practice is using in combination with `legend.justification`.

```{r, legend.position.5, eval = FALSE}
p + theme(
  legend.position = c(0.25, 0),
  legend.justification = c(0, 0) # default "center"
)
```
``` {r, echo = FALSE}
pl <- {
  <<legend.position.5>>
}
wh <- findWH(pl, unit = "in", plot.size = unit(150,"bigpts"), is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE, fig.align='center', out.width="60%"}
pl
```

* a simple illustration shows how it works :

```{r, echo = FALSE, fig.align='center'}
knitr::include_graphics(file.path(getwd(), "04.legend.position.png"))
```

But Comparing to the default layout (left), the legend will no longer be apart of the plot grob tree when using `legend.position` in this way.
The legend will not contribute to panel calculation (see [*panel size*](#panel-size)).

``` {r, echo = F}
gt1 <- ggplotGrob(p)
gt2 <- ggplotGrob(pl)
layout1 <- gtable:::gtable_layout(gt1)
layout2 <- gtable:::gtable_layout(gt2)
```
:::: {.columns style="display: flex;"}
::: {.column width="49%"}
```{r, fig.align='center', fig.width = 10, fig.height = 10, echo = F}
grid:::grid.show.layout(layout1) 
```
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="49%"}
```{r, fig.align='center', fig.width = 10, fig.height = 10, echo = F}
grid:::grid.show.layout(layout2)
```
:::
::::

## legend.box & legend.box.just
`legend.box` controls the arrangement of multiple legends.

`legend.box.just` controls the justification of each legend within the overall bounding box, when there are multiple legends.
```{r, legend.box.1, eval = FALSE}
#default
p + theme(legend.box = "vertical", legend.box.just = "left")
```
```{r, legend.box.2, eval = FALSE}
p + theme(legend.box = "vertical", legend.box.just = "right")
```
```{r, echo = FALSE}
pl <- {
  <<legend.box.2>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE, fig.align = "center", out.width="60%"}
pl
```
```{r, legend.box.3, eval = FALSE}
p + theme(legend.box = "horizontal", legend.box.just = "top")
```
```{r, echo = FALSE}
pl <- {
  <<legend.box.3>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE, fig.align = "center"}
pl
```
```{r, legend.box.4, eval = FALSE}
p + theme(legend.box = "horizontal", legend.box.just = "bottom")
```
```{r, echo = FALSE}
pl <- {
  <<legend.box.4>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE, fig.align = "center"}
pl
```


## legend.box.spacing 
The spacing between the plotting area and the legend box.

See more example in [legend.position](#legend.position), show how it effect in different conditions.

```{r, legend.box.spacing, eval = FALSE}
p + theme(legend.box.spacing = unit(2, "line"))
# default unit(11, "pt")
```
``` {r, echo = FALSE}
pl <- {
  <<legend.box.spacing>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
:::: {.columns style="display: flex;"}
::: {.column width="59%"}
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="39%"}
```{r, fig.dim = wh * c(0.34, 1), echo = FALSE}
show_legend(pl, wh)
```
:::
::::

## legend.box.margin
Margins around the full legend area. 
```{r legend.box.margin, eval = FALSE}
p + theme(legend.box.margin  = margin(2,2,2,2, "line"))
# default margin(0, 0, 0, 0, "cm")
```
``` {r, echo = FALSE}
pl <- {
  <<legend.box.margin>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
:::: {.columns style="display: flex;"}
::: {.column width="59%"}
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="39%"}
```{r, fig.dim = wh * c(0.34, 1), echo = FALSE}
show_legend(pl, wh)
```
:::
::::

## legend.margin
The margin around each legend.
```{r legend.margin, eval = FALSE}
p + theme(legend.margin = margin(2,2,2,2, "line"))
# default margin(5.5, 5.5, 5.5, 5.5, "pt")
```
``` {r, echo = FALSE}
pl <- {
  <<legend.margin>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
:::: {.columns style="display: flex;"}
::: {.column width="59%"}
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="39%"}
```{r, fig.dim = wh * c(0.34, 1), echo = FALSE}
show_legend(pl, wh)
```
:::
::::

## legend.spacing
The spacing between legends.

`legend.spacing.x` controls the <strong>horizontal</strong> gaps between `legend.title` and `legend.key`, between `legend.key` and `legend.text` as well as between multi legends.

`legend.spacing.y` controls the <strong>vertical</strong> gaps between `legend.title` and `legend.key`, between `legend.key` and `legend.text` as well as between multi legends.

```{r legend.spacing.1, eval = FALSE}
pl <- p + guides(
  size  = guide_legend(title = "Size",  ncol = 3, byrow = TRUE),
  alpha = guide_legend(title = "Alpha", ncol = 2)
)
pl + theme(
#   legend.spacing   = NULL, # default unit(11, "pt")
    legend.spacing.x = unit(1, "line"),
    legend.spacing.y = unit(2, "line")
  )
```
``` {r, echo = FALSE}
pl <- {
  <<legend.spacing.1>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
:::: {.columns style="display: flex;"}
::: {.column width="59%"}
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="39%"}
```{r, fig.dim = wh * c(0.34, 1), echo = FALSE}
show_legend(pl, wh)
```
:::
::::

```{r legend.spacing.2, eval = FALSE}
pl + theme(
    legend.spacing.x = unit(1, "line"),
    legend.spacing.y = unit(2, "line"),
    legend.box = "horizontal"
  )
```
``` {r, echo = FALSE}
pl <- {
  <<legend.spacing.2>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
```{r, fig.dim = wh, echo = FALSE}
pl
```


## legend.title & legend.text
```{r legend.title, eval = FALSE}
p + theme(
  legend.title = element_text(size = 15, margin = margin(1,1,1,1, "line")),
  legend.text  = element_text(size = 15, margin = margin(1,1,1,1, "line"))
)
```
``` {r, echo = FALSE}
pl <- {
  <<legend.title>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
:::: {.columns style="display: flex;"}
::: {.column width="59%"}
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="39%"}
```{r, fig.dim = wh * c(0.34, 1), echo = FALSE}
show_legend(pl, wh)
```
:::
::::

## legend.key.size
`legend.key.size` controls the size of legend keys.
```{r legend.key.size.1, eval = FALSE}
p + theme(
    legend.key.size = unit(3, "line") # default unit(1.2, "lines")
#   legend.key.height = NULL,
#   legend.key.width  = NULL
)
```
``` {r, echo = FALSE}
pl <- {
  <<legend.key.size.1>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
:::: {.columns style="display: flex;"}
::: {.column width="59%"}
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="39%"}
```{r, fig.dim = wh * c(0.34, 1), echo = FALSE}
show_legend(pl, wh)
```
:::
::::

`legend.key.size` only effect the box of legend keys instead of the symbol.
To resize the symbol inside legend keys, you need to use `guides`.
<details><summary>example</summary>
```{r legend.key.size.2, eval = FALSE}
p + guides(alpha = guide_legend(title = "Alpha", override.aes = list(size = 20) ))
```
``` {r, echo = FALSE}
pl <- {
  <<legend.key.size.2>>
}
wh <- findWH(pl, unit = "in", is.legend.overlap = TRUE, is.square = TRUE)
```
:::: {.columns style="display: flex;"}
::: {.column width="59%"}
```{r, fig.dim = wh, echo = FALSE}
pl
```
:::
::: {.column width="2%"}
\ 
:::
::: {.column width="39%"}
```{r, fig.dim = wh * c(0.34, 1), echo = FALSE}
show_legend(pl, wh)
```
:::
::::
</details>


## summary
```{r, echo = FALSE, fig.align='center'}
knitr::include_graphics(file.path(getwd(), "04.legend.summary.png"))
```